Examining Programming Students’ Interaction with Generative AI: A Metacognitive Perspective

### 2024-10-04 CSEDU2025/Paper.tex
update paper, related work

@@

About the first research question: \textit{What was the distribution pattern of
the classified messages?}

According to Figure \ref{fig:graph1} messages classified as
Conceptual Question account for 32.1\% of the total. Often, the sequence of
Conceptual Question begins with a request for a Complete Solution or Code
Snippet, which then evolves into a series of Conceptual Question, as seen in
conversation 2 of Figure \ref{fig:graph2}. In other cases, the conversation
consists entirely of messages classified as Conceptual Question, such as in
conversation 10 of Figure \ref{fig:graph2}. This type of message indicates that
students recognize that they have not yet fully mastered a concept and are seeking to
better understand it. In this context, the student realizes that solving the
problem requires understanding of concepts, thus applying an active learning
strategy.

@@

Figure \ref{fig:graph1} also shows that around 5\% of the prompts written by
students were not categorized. These messages range from greetings to the Bot,
such as a simple "\textit{Hola}," to contextual statements like
"\textit{Estoy repasando orientación a objetos.}"

@@

About the second research question: \textit{Was there a space between the student
prompts? If so, which category promotes the largest spacing between
submissions?}

@@

Regarding the third research question: \textit{Do students' conversations with
the bot alternate between practical and theoretical approaches, thereby
promoting interleaving?}

@@

* The results indicate that most interactions are focused on active learning
strategies, although deliberate use of interleaving and spacing is limited. We
believe that students can be guided to adopt a more deliberate approach in
utilizing GenAI, enabling them to more effectively harness these tools to
enhance their learning.

>>>>

\cite{Margulieux24} conducted a study on how undergraduate students in
introductory programming courses used generative AI to solve programming
problems in a naturalistic setting. The research focused on examining the
relationship between AI usage and students’ self-regulation strategies,
self-efficacy, and fear of failure in programming. Additionally, the study
explored how these variables interacted with learner characteristics, perceived
usefulness of AI, and academic performance. Findings revealed that students
with higher self-efficacy, lower fear of failure, or higher prior grades
tended to use AI less frequently or later in the problem-solving process and
perceived it as less useful compared to their peers. However, no significant
relationship was found between students’ self-regulation strategies and their
use of AI.

\cite{Khusnul24} investigates the efficacy of a meta-learning approach in
improving metacognitive and creative skills. This study confirms the
effectiveness of meta-learning strategies and elucidates the relationship
between meta-learning and metacognition. This research suggests that
meta-learning can improve metacognitive abilities, providing valuable insights
into educational technology and course design in higher education settings.

Reconceptualizing AI Literacy: The Importance of Metacognitive Thinking in an Artificial Intelligence (AI)-Enabled Workforce

@@

About the first research question: \textit{RQ1: What was the distribution
pattern of the classified messages?}

According to Figure \ref{fig:graph1} messages classified as
Conceptual Question account for 32.1\% of the total. Often, the sequence of
Conceptual Question begins with a request for a Complete Solution or Code
Snippet, which then evolves into a series of Conceptual Question, as seen in
conversation 2 of Figure \ref{fig:graph2}, which illustrates some examples of
students' conversations with the bot. In other cases, the conversation
consists entirely of messages classified as Conceptual Question, such as in
conversation 10 of Figure \ref{fig:graph2}. This type of message indicates that
students recognize that they have not yet fully mastered a concept and are
seeking to better understand it. In this context, the student realizes that
solving the problem requires understanding of concepts, thus applying an active
learning strategy.

@@

Figure \ref{fig:graph1} also shows that around 5\% of the prompts written by
students were not categorized. These messages range from greetings to the Bot,
such as a simple "\textit{Hello}," to contextual statements like
"\textit{I'm reviewing object orientation}"

@@

About the second research question: \textit{RQ2: Was there a space between the
student prompts? If so, which category promotes the largest spacing between
submissions?}

@@

Regarding the third research question: \textit{RQ3: Do students' conversations
with the bot alternate between practical and theoretical approaches, thereby
promoting interleaving?}

@@

* The results indicate that most interactions are focused on active learning
strategies, although deliberate use of interleaving and spacing is limited. We
believe that students can be guided \cite{Zheng19} \cite{Wang23} \cite{LiWei23}
\cite{Khusnul24} to adopt a more deliberate approach in utilizing GenAI,
enabling them to more effectively harness these tools to enhance their learning.
